---
title: 'Selective Dry Cow Therapy: Quarter-level analysis descriptive'
output:
  html_notebook:
    toc: true
    toc_float:
      collapsed: false

---

```{r}
library(knitr)
```

```{r setup}
#Mac
knitr::opts_knit$set(root.dir = "~/Dropbox/R backup/SDCT - R")

#Windows
#knitr::opts_knit$set(root.dir = "C:/Users/rowe0122/Dropbox/R backup/SDCT - R")

```


```{r}
library(knitr)
```

Load data
```{r}
load(file="SDCTQTR.Rdata")
```


# Evaluate Data
## Show first 10 rows of baseline
```{r}
load(file="Baseline.Rdata")
head(Baseline, n=10)
```

<br>
<br>
## Inspect Data 
```{r}
#summarytools::dfSummary(Baseline, style='grid')
print(summarytools::dfSummary(Baseline, valid.col=FALSE, graph.magnif=0.8, style="grid"), method = "render")
```
<br>
<br>
## Pathogens
```{r}
table1::table1(~ DOCult2 | Tx, data=SDCTQTR)
table1::table1(~ PCCult1 | FARMID, data=SDCTQTR)
```


## Compare treatment groups at baseline
```{r}
table1::table1(~ factor(FARMID) + Spectramast + Age + Parity + factor(IMIDO) + DOSCC + DOMY + DODIM + PrevCM + PrevSCCHI + DPlength + PCSampDIM | Tx, data=SDCTQTR)
```

<br>
<br>
<br>
## Compare herds
```{r}
table1::table1(~ Age + Parity  + DOSCC + DOMY + DODIM + PrevCM + PrevSCCHI + DPlength + PCSampDIM + factor(IMIDO) + factor(IMIPC) + factor(Cure) + factor(NewIMI) | factor(FARMID), data=SDCTQTR)
```

# Outcome 1: Cure
```{r}
CrossTable(SDCTQTR$Tx,SDCTQTR$Cure,prop.c=FALSE,prop.t=FALSE,prop.chisq = FALSE)
```

## Logistic regression model for dry period cure
## Model building plan

Model type: Logistic regression with mixed effects (generalized linear mixed model with binomial family / logit link).  

Step 1: Identify potential confouders using a directed acyclic graph (DAG)

Step 2: Identify correlated variables using pearson and kendalls correlation coefficients

Step 3: Create model with all potential confounders

Step 4: Investigate potential effect measure modification

Step 5: Remove unneccesary covariates in backwards stepwise fashion using 10% rule (i.e. if odds ratio for algorithm or culture changes by >10% after removing the covariate, the covariate is retained in the model)

Step 6: Report final model

### Step 1: DAG
This is used to identify variables that could be confounders if they are not balanced between treatment groups. 
```{r fig.width = 10}
library(DiagrammeR)
mermaid("graph LR
        T(Treatment)-->U(Cure)
        A(Age)-->T
        P(Parity)-->T
        M(Yield at dry-off)-->T
        S(SCC during prev lactation)-->T
        C(CM in prev lact)-->T
        D(DIM at dry off) --> T
        K(DIM at post calving sample) --> T
        D-->M
        D-->S
        D-->U
        K-->U
        A-->U
        P-->U
        M-->U
        S-->U
        C-->U
        C-->M
        P-->C
        P-->S
        P-->M
        A-->P
        A-->C
        A-->S
        A-->M
        M-->S
        C-->S
style D fill:#FFFFFF, stroke-width:0px
style K fill:#FFFFFF, stroke-width:0px
style A fill:#FFFFFF, stroke-width:0px
        style T fill:#FFFFFF, stroke-width:2px
        style P fill:#FFFFFF, stroke-width:0px
        style M fill:#FFFFFF, stroke-width:0px
        style S fill:#FFFFFF, stroke-width:0px
        style C fill:#FFFFFF, stroke-width:0px
        style I fill:#FFFFFF, stroke-width:0px
        style U fill:#FFFFFF, stroke-width:2px
        ")
```
According to this DAG, I may need to control for the following variables.

Parity ["Parity"] or Age ["Age"] <- likely to correlated

Yield at most recent test before dry off ["DOMY"]

Somatic cell count at last herd test during previous lactation ["DOSCC" or "PrevSCCHi"] <- likely to be correlated

Clinical mastitis in previous lactation ["PrevCM"]

Days in milk at dry-off ["DODIM"]

Days in milk at post-calving sample ["PCSampDIM"]

<br>
<br>
<br>
<br>

### Step 2: Identify correlated covariates
Pearson correlation matrix among potential predictors
```{r}
cor <- Baseline[, c(6,7,9,10,11,13,14,15)]
cor$Parity <- as.numeric(cor$Parity)
#corrplot(cor)
cor <- cor(cor, use = "complete.obs")
round(cor, 2)
```

```{r}
cor <- Baseline[, c(6,7,9,10,11,13,14,15)]
cor$Parity <- as.numeric(cor$Parity)
cor <- cor(cor, use = "complete.obs", method="kendal")
round(cor, 2)
```

Age and Parity highly correlated as expected.  Will only offer parity. 

Previous lactation peak SCC is moderately correlated with SCC at last herd test (DOSCC). Will only offer DOSCC. 

<br>
<br>
<br>
<br>


# Outcome 2: IMI at Calving
```{r}
CrossTable(SDCTQTR$Tx,SDCTQTR$IMIPC,prop.c=FALSE,prop.t=FALSE,prop.chisq = FALSE)
```
## Logistic regression model for IMI at calving
## Model building plan

Model type: Logistic regression with mixed effects (generalized linear mixed model with binomial family / logit link).  

Step 1: Identify potential confouders using a directed acyclic graph (DAG)

Step 2: Create model with all potential confounders

Step 3: Investigate potential effect measure modification

Step 4: Remove unneccesary covariates in backwards stepwise fashion using 10% rule (i.e. if odds ratio for algorithm or culture changes by >10% after removing the covariate, the covariate is retained in the model)

Step 5: Report final model

### Step 1: DAG
This is used to identify variables that could be confounders if they are not balanced between treatment groups. 
```{r fig.width = 10}
library(DiagrammeR)
mermaid("graph LR
        T(Treatment)-->U(IMI at calving)
        A(Age)-->T
        P(Parity)-->T
        M(Yield at dry-off)-->T
        S(SCC during prev lactation)-->T
        C(CM in prev lact)-->T
        D(DIM at dry off) --> M
        I(IMI at dry off) --> T
        K(DIM at post calving sample) --> T
        I-->U
        P-->I
        A-->I
        C-->I
        I-->S
        I-->M
        D-->S
        D-->U
        K-->U
        A-->U
        P-->U
        M-->U
        S-->U
        C-->U
        C-->M
        P-->C
        P-->S
        P-->M
        A-->P
        A-->C
        A-->S
        A-->M
        M-->S
        C-->S
style D fill:#FFFFFF, stroke-width:0px
style K fill:#FFFFFF, stroke-width:0px
style A fill:#FFFFFF, stroke-width:0px
        style T fill:#FFFFFF, stroke-width:2px
        style P fill:#FFFFFF, stroke-width:0px
        style M fill:#FFFFFF, stroke-width:0px
        style S fill:#FFFFFF, stroke-width:0px
        style C fill:#FFFFFF, stroke-width:0px
        style I fill:#FFFFFF, stroke-width:0px
        style U fill:#FFFFFF, stroke-width:2px
        ")
```
According to this DAG, I may need to control for the following variables.

Parity ["Parity"] or Age ["Age"] <- will use Parity

Yield at most recent test before dry off ["DOMY"]

Somatic cell count at last herd test during previous lactation ["DOSCC" or "PrevSCCHi"] <- will use DOSCC

Clinical mastitis in previous lactation ["PrevCM"]

IMI at dry-off ["DOIMI"]

Days in milk at dry-off ["DODIM"]

Days in milk at post-calving sample ["PCSampDIM"]

<br>
<br>
<br>
<br>



# Outcome 3: New IMI
```{r}
CrossTable(SDCTQTR$Tx,SDCTQTR$NewIMI,prop.c=FALSE,prop.t=FALSE,prop.chisq = FALSE)
```

## Logistic regression model for new IMI
## Model building plan

Model type: Logistic regression with mixed effects (generalized linear mixed model with binomial family / logit link).  

Step 1: Identify potential confouders using a directed acyclic graph (DAG)

Step 2: Create model with all potential confounders

Step 3: Investigate potential effect measure modification

Step 4: Remove unneccesary covariates in backwards stepwise fashion using 10% rule (i.e. if odds ratio for algorithm or culture changes by >10% after removing the covariate, the covariate is retained in the model)

Step 5: Report final model

### Step 1: DAG
This is used to identify variables that could be confounders if they are not balanced between treatment groups. 
```{r fig.width = 10}
library(DiagrammeR)
mermaid("graph LR
        T(Treatment)-->U(New IMI)
        A(Age)-->T
        P(Parity)-->T
        M(Yield at dry-off)-->T
        S(SCC during prev lactation)-->T
        C(CM in prev lact)-->T
        D(DIM at dry off) --> T
        I(IMI at dry off) --> T
        K(DIM at post calving sample) --> T
        I-->U
        P-->I
        A-->I
        C-->I
        I-->S
        I-->M
        D-->S
        D-->M        
        D-->U
        K-->U
        A-->U
        P-->U
        M-->U
        S-->U
        C-->U
        C-->M
        P-->C
        P-->S
        P-->M
        A-->P
        A-->C
        A-->S
        A-->M
        M-->S
        C-->S
style D fill:#FFFFFF, stroke-width:0px
style K fill:#FFFFFF, stroke-width:0px
style A fill:#FFFFFF, stroke-width:0px
        style T fill:#FFFFFF, stroke-width:2px
        style P fill:#FFFFFF, stroke-width:0px
        style M fill:#FFFFFF, stroke-width:0px
        style S fill:#FFFFFF, stroke-width:0px
        style C fill:#FFFFFF, stroke-width:0px
        style I fill:#FFFFFF, stroke-width:0px
        style U fill:#FFFFFF, stroke-width:2px
        ")
```
According to this DAG, I may need to control for the following variables.

Parity ["Parity"] or Age ["Age"] <- will use Parity

Yield at most recent test before dry off ["DOMY"]

Somatic cell count at last herd test during previous lactation ["DOSCC" or "PrevSCCHi"] <- will use DOSCC

Clinical mastitis in previous lactation ["PrevCM"]

IMI at dry-off ["DOIMI"]

Days in milk at dry-off ["DODIM"]

Days in milk at post-calving sample ["PCSampDIM"]

<br>
<br>
<br>
<br>
